\chapter{About APECSS}

APECSS is a software toolbox to compute pressure-driven bubble dynamics and the resulting acoustic emissions. It is written exclusively in {\tt C} and has been developed with simplicity, versatility and performance in mind. The acronym APECSS stands for "Acoustic Pulse Emitted by Cavitation in Spherical Symmetry".

The main features of APECSS are:\vspace{-1em}
\begin{itemize}[noitemsep]
\item Bubble dynamics using widely-used models (Rayleigh-Plesset, Keller-Miksis, Gilmore).
\item Acoustic emissions of the bubble under different assumptions (incompressible, quasi-acoustic, fully compressible).
\item Acoustic interactions of multiple bubbles under different assumptions (incompressible, quasi-acoustic)
\item Prediction of the formation and attenuation of shock fronts emitted by the bubble.
\item Viscoelastic media (Kelvin-Voigt, Zener, Oldroyd-B).
\item Lipid monolayer coating of the bubble as used for ultrasound contrast agents.
\item All ODEs are solved with in-built fourth- and fifth-order Runge-Kutta scheme with adaptive time stepping.
\item APECSS has, aside from the C standard library, no external dependencies.
\end{itemize}

The APECSS repository is located at \href{https://github.com/polycfd/apecss}{\texttt{https://github.com/polycfd/apecss}} and structured as:\vspace{-1em}
\begin{itemize}[noitemsep]
    \item The {\tt documentation/} folder contains this short documentation of APECSS.
    \item The {\tt examples/} folder contains representative examples of how to use APECSS and to demonstrate the most important features of APECSS. 
    \item The {\tt include/} folder contains the {\tt apecss.h} header file, in which all variables, macros and functions of APECSS are defined.
    \item The {\tt lib/} folder in which the APECSS library is compiled (at least if you follow the installation instructions in Section \ref{sec:installation}).
    \item The {\tt src/} folder contains all source files ({\tt *.c}) of APECSS.
    \item The {\tt .clang-format} file, which defines the formatting rules for the source code, see Section \ref{sec:clang}.
    \item The {\tt .gitignore} file tells \textit{git} which folders and files to ignore.
    \item The {\tt CITATION.cff} file contains information on how to cite APECSS.
    \item The {\tt CODE\_OF\_CONDUCT.md} file outlines the code of conduct for this repository.
    \item The {\tt CONTRIBUTING.md} file contains brief guidelines on how to contribute to APECSS.
    \item The {\tt JOSS-Paper.pdf} file is the paper about APECSS in the Journal of Open Source Software.
    \item The {\tt LICENSE.txt} file contains the Mozilla Public License Version 2.0.
    \item The {\tt README.md} file with the most important information about APECSS.
\end{itemize}

APECSS is under the copyright of its developers and made available as open-source software under the terms of the \href{https://www.mozilla.org/en-US/MPL/2.0/}{Mozilla Public License Version 2.0}.

The development of APECSS has directly benefitted from research funding provided by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation), grant number 441063377, and by the Natural Sciences and Engineering Research Council of Canada (NSERC), funding reference number RGPIN-2024-04805.
